# Rstudio
This chapter contains syntax for manipulating data and packages within the  
R studio environment.

## Useful packages
Load all libraries
```{r}
library(tidyverse) # data handling and viz
library(janitor) #dataframe import cleaning
library(knitr) #nice html tables
library(kableExtra) # nicer knitr tables
library(broom)
library(readr) # load csv stored data
library(geosphere) # for calc daylength
```
## Remove a package
```{r}
#Unload a module: 
library(clipr) #load
detach(package:clipr) #unload
```

## Import using Janitor
```{r}
# Create a data.frame with dirty names
test_df <- as.data.frame(matrix(ncol = 6))
names(test_df) <- c("firstName", "ábc@!*", "% successful (2009)",
                    "REPEAT VALUE", "REPEAT VALUE", "")
head(test_df)
test_df <- test_df %>%
  clean_names()
head(test_df)
```
[Reference](https://cran.r-project.org/web/packages/janitor/vignettes/janitor.html)

## Remove dataframe 
```{r}
data("mtcars")
data("band_instruments")
data("band_instruments2") # Load example datasets

rm(list=ls()[! ls() %in% c("band_instruments","band_instruments2")]) 
# Everything except Band instruments
rm(list=setdiff(ls(), "band_instruments")) # Everything except "bandinstruments"
rm(list=ls()) # Remove everything
```
[Reference:Stackoverflow](https://stackoverflow.com/questions/6190051/how-can-i-remove-all-objects-but-one-from-the-workspace-in-r)

## Save dataframe  (CSV or Rdata)
*make date string*
```{r}
datenow <- format(Sys.time(), "_%Y_%m_%d")
date
data(mtcars)
```
*Write file names*
```{r}
#create data directory
dir.create("data_out")

filenamecsv <- paste("data_out/mtcsvdata",datenow,".csv",sep="")
filenamerda <- paste("data_out/mtrdadata",datenow,".rda",sep="")
```
*Save the files*
```{r}
save(mtcars,file=filenamerda)
write.csv(mtcars,file=filenamecsv)
```

## Save a diagram or plot
```{r}
plot1 <- mtcars %>% ggplot(aes(hp,qsec))+geom_point()
#plot1 #print plot if required
pdf("plot.pdf")
plot1
dev.off()
```
pdf 'device' off.

## Recode a text variable
```{r message=FALSE}
data("band_members")
kable(head(band_members)) %>% kable_minimal(full_width = F)
band_members <- band_members %>% mutate(name=recode(name, "Mick"= "m"))
kable(head((band_members))) %>% kable_minimal(full_width = F)
rm(list=ls()) # Remove everything
```
Reference: [Kable Extra](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

## Alter variable names:
*Remove underscores*
```{r}
data("mtcars")
mtcars <- mtcars %>% rename(hp_new=hp)
kable(head((mtcars))) %>% kable_minimal(full_width = F)

mtcars <- mtcars %>% rename_with(.fn = ~str_replace(., "_", ""))
kable(head((mtcars))) %>% kable_minimal(full_width = F)
```

### list datasets available
```{r}
#data() # list all available datasets
data("diamonds")
```
