# Data Wrangling and manipulation

```{r}
library(Hmisc) #impute values
library(naniar) # deal with NAs
library(geosphere)
library(tidyverse) # data handling and viz
library(janitor) #dataframe import cleaning
library(knitr) #nice html tables
library(kableExtra) # nicer knitr tables
library(broom)
library(readr) # load csv stored data
library(geosphere) # for calc daylength
```


## Bin variable  ( e.g. Low/Medium/High)
```{r}
data(mtcars)
mtcars <- mtcars %>% mutate(hp_cat=cut(hp, breaks=c(-Inf, 100, Inf), 
                                       labels=c("low hp","high hp")))
```

## Conditional function
```{r}
mtcars <- mtcars %>% mutate(loghp=ifelse(cyl>4,log10(hp),NA)) 
# Nonsensical example, but log transformed all horse powers of cars with more
# than four cylinders
```

## Sum across rows
```{r}
mtcars <- mtcars %>%  mutate(sum = select(., disp:drat) %>% 
apply(1, sum, na.rm=TRUE))
#apply() takes Data frame or matrix as an input and gives output in vector 
#(i.e.many columns to one list)
# the '1' sets the dataframe to use (already selected here)

```
[Reference](https://stackoverflow.com/questions/28873057/sum-across-multiple-columns-with-dplyr)

## Standardise variable 
```{r stardardise variable 1}
dat2 <- mtcars %>%
    as_tibble() %>%
    mutate(across(where(is.numeric), scale))

funcs <- list(mean = ~mean(.x,na.rm = TRUE), 
  sd = ~sd(.x,na.rm = TRUE)
)
dat2 %>% summarise(across(where(is.numeric),funcs))
```
## Conditional Replacement
*Replace all 'NA's in a specified variable with 0.*
```{r}
mtcars <- mtcars %>% mutate(loghp1 = coalesce(loghp, 0))
#or
mtcars <- mtcars %>% mutate(loghp = replace_na(loghp, "missing"))
```

## Filter na's or retain complete cases
```{r}
mtcars <- mtcars %>% filter(!is.na(hp)) # no missing values found
mtcars <- mtcars %>%filter(complete.cases(.)) # no missing values found
```

## Delete specified columns
```{r}
mtcars1 <- mtcars %>% select(-(drat)) # single column
mtcars2 <- mtcars %>% select(-c(drat,hp,vs:gear)) # multiple columns

rm(list=setdiff(ls(), "mtcars")) # clean environment
```

##Change specific datapoint

```{r}
mtcarsmissingvalues <- mtcars %>% mutate(gear=ifelse(gear==5,"missing",gear))
```

## Find duplicate rows
```{r}
# specify which variable to check for duplication
n_occur1 <- data.frame(table(mtcars$mpg)) 
kable(n_occur1[n_occur1$Freq > 1,]) %>% kable_styling(full_width = F) %>% kable_minimal()
```

## Impute missing values
### To be completedImputing missing values using the mean:

```{r}
#create missing values
#mtcarsmissingvalues <- mtcars %>% mutate(gear=ifelse(gear==5,"",gear))

mtcarsmissingvalues <- mtcars %>% replace_with_na(replace = list(gear = 5))
mtcarsmissingvalues$gear <- impute(mtcarsmissingvalues$gear, mean)  # replace with mean
mtcarsmissingvalues$gear <- impute(mtcarsmissingvalues$gear, median) # median
mtcarsmissingvalues$gear <- impute(mtcarsmissingvalues$gear, 4) # replace specific number

```

[Reference:](http://r-statistics.co/Missing-Value-Treatment-With-R.html)


## Keep rows based on a unique value.
e.g. prescription code
```{r}
mtcarsdistinct <- mtcars %>% 	distinct(cyl, .keep_all= TRUE)
```
[Reference](http://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function)

## Delete rows on a variable value
```{r}
mtcars1<-mtcars %>% filter(!(cyl==6))
mtcars2<-mtcars %>% filter(!(cyl==6 | hp==180)) # | is the 'or' operator
mtcars3<-mtcars %>% filter(!(cyl==8 & hp==215)) # & is the 'and' operator
# remove the ! To select the individuals with the specified conditions
```

## Use if else to calculate  on values
```{r}
# no NA's so all values unchanged.
mtcars <- mtcars %>% mutate(vs=ifelse(is.na(vs),(carb-am)/365.25,vs)) 
```

## Merge data frames (variables)
*left_join(x, y): returns all rows from x, and all columns from x and y. Rows in   x with no match in y will have NA values in the new columns. If there are multiple matches between x and y, all   combinations of the matches are returned.

*inner_join(x, y): returns all rows from x where there are matching values in y,   and all columns from x and y. If there are multiple matches between x and y, all combinations of the matches are returned.

*full_join(x, y): returns all rows and all columns from both x and y. Where there are not matching values, the function returns NA for the one missing

* inner: only rows with matching keys in both x and y
* left: all rows in x, adding matching columns from y
* right: all rows in y, adding matching columns from x
* full: all rows in x with matching columns in y, then the rows of y that don't match x.

```{r}
# prepare new dataset
# make the rownames into a 'joinable' column
mtcars <- mtcars %>% mutate(carnames=rownames(mtcars)) 
mtcars_extradata <- mtcars %>% select(cyl)
# make the rownames into a 'joinable' column
mtcars_extradata <- mtcars_extradata %>%
mutate(carnames=rownames(mtcars_extradata)) 
mtcars_extradata <- mtcars_extradata %>% mutate(valves=cyl*4)
mtcars_extradata <- mtcars_extradata %>% select(-cyl)

kable(glimpse(mtcars_extradata%>% slice(1:6))) %>%  
  kable_styling(full_width = F) %>%
  kable_minimal()

mtcars <- left_join(mtcars,mtcars_extradata,by = 'carnames')

kable(glimpse(mtcars %>%select(carb:valves) %>%  slice(1:6))) %>% kable_styling(full_width = F) %>%
  kable_minimal()

```

## Merge data frames (individuals)
```{r}
mtcarsmerged <- bind_rows(mtcars2, mtcars3)
rm(list=setdiff(ls(), "mtcars")) # clean environment
```
[Reference](https://dplyr.tidyverse.org/reference/bind.html)

## Create a new factor from existing 
```{r}
mtcars <- mtcars %>% mutate(cyc_carb = paste(cyl,carb,sep="-"))
```

## change data types
(merging fails if data types are different)
```{r}
# adni_demog<-adni_demog %>% mutate(age_scan=as.numeric(age_scan))
# ukbb<-ukbb %>% mutate(scan_no=as.numeric(scan_no))
```

## calculate dates and photoperiod 
(using geosphere library)
```{r}
#import sample dataset
dateslat <- read_csv("dateslat.csv")
dateslat <- dateslat %>%
  clean_names()

dateslat <- dateslat %>% mutate(dateofscan=(as.Date(date_dmy,format="%d/%m/%Y")))
dateslat <- dateslat %>% mutate(daylength=daylength(latitude,dateofscan))
                                
dateslat %>% ggplot(aes(x=dateofscan,y=daylength)) +geom_line() +theme_minimal()
```

## Reduce variables using PCA
### To be completed
