# Data Visualisation
**Tables and graphs, survival plots, missing values.**

## Packages needed
```{r include=FALSE}

library(arsenal) # for summary tables
library(summarytools) # for summary tables
library(gridExtra) # print multiple plots as grid
library(ggpmisc) # add formulas and p values to scatterplots
library(corrplot) #plotting correlations
library(Hmisc) #impute values
library(naniar) # deal with NAs
library(geosphere)
library(tidyverse) # data handling and viz
library(janitor) #dataframe import cleaning
library(knitr) #nice html tables
library(kableExtra) # nicer knitr tables
library(broom)
library(readr) # load csv stored data
library(geosphere) # for calc daylength
#library(RColorBrewer)
library(viridis)
library(reshape2)
library(ggrepel) # label points on ggplot
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
          
```
## Summary Tables
### Summarise by group
```{r}
data(mtcars)
kable(mtcars %>% group_by(cyl) %>% summarise(Ave=mean(hp), StDev=sd(hp))) %>% 
  kable_styling(full_width = FALSE) %>% kable_minimal()
```

###  Summary Table - Multiple functions, variables
```{r}
# make sure brackets are correct

df.sum <- mtcars %>%  select(mpg,cyl,hp) %>%
  summarise(across(everything(),list(mean=mean,sd=sd)))
kable(df.sum,digits=2) %>% kable_styling(full_width = FALSE) %>%
  kable_minimal() # perform the analysis

df.longer <- df.sum%>% pivot_longer(col=everything(),  
names_to = c("Attribute",".value"),
  names_sep = "_")
kable(df.longer,digits=2) %>%
  kable_styling(full_width = FALSE) %>%
  kable_minimal() # pivot longer the analysis to make it readable
```

### 'Arsenal' summary table


```{r, results='asis'}
tab1 <- tableby(cyl~gear+hp+wt,data=mtcars)
summary(tab1, text=TRUE, digits=2, digits.p=2, digits.pct=1)
```
### Summarytools tables
```{r}
descr(mtcars, stats = c("mean", "sd"), transpose = TRUE, headings = FALSE)

kable(descr(mtcars, stats = c("mean", "sd", "n.valid"), transpose = TRUE,
              headings = FALSE),digits = 3) %>% 
  kable_styling(full_width = FALSE)%>% kable_minimal()
```

## Visual summary of data
*Options are for markdown*
```{r, results='asis'}

dfSummary(mtcars, plain.ascii = FALSE, style = "grid", 
          graph.magnif = 0.5, valid.col = FALSE, tmp.img.dir = "/tmp")
```
## Correlation matrix
### Ellipse style
```{r, fig.cap="Correlation Plot"}
corrdata <- mtcars %>% select(-c(cyl,disp,vs,am,gear,carb)) 
corr1 <- Hmisc::rcorr(as.matrix(corrdata))
M <- corr1$r
#M
colnames(M) <- c("mpg", "HP", "Axle Ratio", "Weight (kPounds)", "Quarter Mile (s)")
rownames(M) <-  c("mpg", "HP", "Axle Ratio", "Weight (kPounds)", "Quarter Mile (s)")
p_mat <- corr1$P
corr <- corrplot(M, type = "upper",method="ellipse", order = "hclust", 
         p.mat = p_mat, sig.level = 0.05, insig = "blank")
```

* Red is -ve correlation
* Blue is + ve correlation
* Blank is no correlation

[Reference](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

## Graphing
### Frequency Histogram - basic
```{r}
plot1 <- mtcars %>% ggplot(aes(qsec))+geom_histogram()
plot1
```
### Frequency Histogram + normal distribution
```{r}
plot1 <- mtcars %>% ggplot(aes(qsec))
#plot1+geom_histogram()
# add normal plot
plot1 + geom_histogram(aes( y=..density..))+
  stat_function(fun = dnorm, args = list(mean =mean(mtcars$qsec), sd=sd(mtcars$qsec))) + 
  theme_bw()
# ..density.. changes y axis to density, not count. stat function defines normal
# line based on data provided.
```

### multiple plot of all distributions
```{r}
mtcars %>% keep(is.numeric) %>% gather() %>% ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") + geom_histogram()
```

### x*y scatterplot with  linear or polynomial regression
```{r echo=TRUE}
plot2<- mtcars %>% ggplot(aes(x=wt,y=qsec))
plot2a <- plot2 +geom_point()+stat_smooth(method='lm',formula=y~x) + theme_bw()
plot2b <- plot2 +geom_point()+stat_smooth(method='lm',formula = y ~ poly(x, 2)) + theme_bw()
plot2c <- plot2 +geom_point()+stat_smooth(method='lm',formula = y ~ poly(x, 3)) + theme_bw()
plot2d <- plot2 +geom_point()+stat_smooth(method='lm',formula = y ~ poly(x, 4)) + theme_bw()

grid.arrange(plot2a,plot2b,plot2c,plot2d,nrow=2,ncol=2)
```

### Add formula to plot.
```{r}
my.formula <- y ~ x
a <- plot2 +geom_point()+geom_smooth(method='lm',formula=my.formula)+
  stat_poly_eq(formula = my.formula, aes(label = paste(..eq.label.., ..rr.label..,
                                                       sep = "~~~")), parse = TRUE) +
  theme_bw()

my.formula2 <- y ~ poly(x, 2)
b <- plot2 +geom_point()+geom_smooth(method='lm',formula=my.formula2)+
  stat_poly_eq(formula = my.formula2, aes(label = paste(..eq.label.., ..rr.label..,
                                                        sep = "~~~")), parse = TRUE) +
  theme_bw()

grid.arrange(a,b,nrow=1)
```

### Raincloud plots (ggplot)
```{r}
library(plyr)
library(dplyr)
```

#### custom theme creation
```{r}
theme_rain = theme(
text = element_text(size = 10),
axis.title.x = element_text(size = 16),
axis.title.y = element_text(size = 16),
axis.text = element_text(size = 14),
axis.text.x = element_text(angle = 0, vjust = 0.5),
legend.title=element_text(size=16),
legend.text=element_text(size=16),
legend.position = "right",
plot.title = element_text(lineheight=.8, face="bold", size = 16),
panel.border = element_blank(),
panel.grid.minor = element_blank(),
panel.grid.major = element_blank(),
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))
```

#### make summary functions
```{r}
lb <- function(x) mean(x) - sd(x)
ub <- function(x) mean(x) + sd(x)
```

#### row names as real column
```{r}
mtcars <- tibble::rownames_to_column(mtcars, "car_name")
mtcars <- mtcars %>% mutate(cyl=as_factor(cyl))
```
#### calc summary data
```{r}
data("diamonds")
sumld<- ddply(diamonds, ~cut, summarise, mean = mean(carat), median = median(carat), 
              lower = lb(carat), upper = ub(carat))
kable(head(sumld)) %>% kable_minimal()
```

### raincloud plot (diamonds)
```{r}


g <- ggplot(data = diamonds, aes(y = carat, x = cut, fill = cut)) +
geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .3) +
geom_point(aes(y = carat, color = cut), position = position_jitter(width = .15), size = .5, alpha = 0.3) +
geom_boxplot(width = .2, guides = FALSE, outlier.shape = NA, alpha = 0.9) +
expand_limits(x = 5.25) +
scale_color_viridis_d() +
scale_fill_viridis_d() +
coord_flip() +
theme_bw() +
theme_rain

g

```
### Alternative raincloud

```{r, fig.cap="Raincoud plot of means"}
#calculations needed
sumld<- ddply(diamonds, ~cut, summarise, mean = mean(carat), median = median(carat), lower = lb(carat), upper = ub(carat))

g <- ggplot(data = diamonds, aes(y = carat, x = cut, fill = cut)) +
geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8) +
geom_point(aes(y = carat, color = cut), position = position_jitter(width = .15), size = .5, alpha = 0.8) +
geom_point(data = sumld, aes(x = cut, y = mean), position = position_nudge(x = 0.3), size = 2.5) +
geom_errorbar(data = sumld, aes(ymin = lower, ymax = upper, y = mean), position = position_nudge(x = 0.3), width = 0) +
expand_limits(x = 5.25) +
guides(fill = FALSE) +
guides(color = FALSE) +
scale_color_viridis_d() +
scale_fill_viridis_d() +
theme_bw() +
theme_rain

g
```

### Scatterplot theme
```{r}
theme_scatter = theme(
text = element_text(size = 10),
axis.title.x = element_text(size = 12),
axis.title.y = element_text(size = 12, angle = 0,vjust = .5),
axis.text = element_text(size = 10),
axis.text.x = element_text(angle = 0, vjust = 0.5),
legend.title=element_text(size=12,hjust = .5),
legend.text=element_text(size=10),
#legend.position = "right",
legend.background = element_rect(colour='light grey'),
plot.title = element_text(lineheight=.8, face="bold", size = 16),
panel.border = element_blank(),
panel.grid.minor = element_blank(),
panel.grid.major = element_blank(),
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))
```



### Scatterplots
```{r, fig.cap="Scatterplots"}

sp <- diamonds %>% ggplot(aes(x=carat,y=price))
sp1 <- sp+geom_point()
sp2 <- sp+geom_point() +theme_bw()
sp3 <- sp+geom_point() +theme_bw() +theme_scatter
sp4 <- sp+geom_point(alpha=.01)+ylab('(£)') +theme_bw() +theme_scatter


grid.arrange(sp1,sp2,sp3,sp4,nrow=2,ncol=2)
```

### make axis logarithmic
```{r}
sp5 <- sp+geom_point(alpha=.01)+ylab('(£)') +theme_bw() +theme_scatter
sp5+ scale_x_continuous(trans='log10') +
  scale_y_continuous(trans='log10')
```

### add a trendline
```{r, fig.cap='linear Trendline'}
my.formula <- y ~ x # calc formula for display

sp5+ylim(0,20000)+xlim(0,3)+geom_smooth(method='lm',formula =my.formula,
                                        colour='black', size=.4,alpha=.6)+
  stat_poly_eq(formula = my.formula,
               aes(label = paste(..eq.label.., ..rr.label..,
                                                        sep = "~~~")), parse = TRUE)
```

### add a trendline
```{r, fig.cap='Polynomial trendline'}
formula <- y ~ poly(x, 2,raw=TRUE) # calc formula for display

sp5+ylim(0,20000)+xlim(0,3)+geom_smooth(method='lm',formula =formula,
                                        colour='black', size=.4,alpha=.6)+
  stat_poly_eq(formula = formula,
               aes(label = paste(..eq.label.., ..rr.label..,
                                                        sep = "~~~")), parse = TRUE)
```

### add multiple trendlines
```{r message=FALSE, warning=FALSE}
sp <- diamonds %>% ggplot(aes(x=carat,y=price,colour=clarity))
sp5 <- sp+geom_point(alpha=.01)+ylab('(£)') +theme_bw() +theme_scatter
sp6 <- sp5+ylim(0,20000)+xlim(0,3)+ guides(colour = guide_legend(override.aes = list(alpha = 1)))
my.formula4 <- y ~ poly(x, 2,raw = TRUE) # calc formula for display

sp7 <- sp6+geom_smooth(aes(colour=clarity),method='lm',formula = my.formula4,se=F, size=.4,alpha=.6)
sp7

sp8 <- sp7+
  stat_poly_eq(aes(label =  paste(stat(eq.label),
                                  stat(adj.rr.label), sep = "*\", \"*")),
               formula = my.formula4, parse = TRUE,size=3,  geom = "label_npc")
sp8

```
### add deviation from regression
```{r,}
data(mtcars)
mtcars <- tibble::rownames_to_column(mtcars, "car_name")
```


```{r, fig.cap='Deviation from predection'}
formula <- y ~ poly(x, 2,raw=TRUE) # calc formula for display

hpvmpg <- mtcars %>% ggplot(aes(x=hp,y=mpg,label=rownames(mtcars)))
# p1 <- hpvmpg+geom_point()+geom_smooth(method='lm',formula =formula,colour='black', size=.4,alpha=.6,se=F)+
#   stat_fit_deviations(formula = formula, colour = "red")+geom_label_repel(aes(label =rownames(mtcars)),
#                   box.padding   = 0.1, 
#                   point.padding = 0.3,
#                   segment.color = 'grey50')

p2 <- hpvmpg+geom_point()+geom_smooth(method='lm',formula =formula,colour='black', size=.4,alpha=.6,se=F)+
  stat_fit_deviations(formula = formula, colour = "red")+geom_label_repel(aes(label =rownames(mtcars)),
    arrow = arrow(length = unit(0.02, "npc")),
    box.padding = .5,min.segment.length = 0,max.overlaps = Inf)

 p2
 
#grid.arrange(p1,p2)
```
### residuals (ID those >1SD from $bar{X})
```{r}
lm <- lm(hp ~ poly(disp, 2,raw=TRUE),data=mtcars) # make lin model
resids <- resid(lm) # extract resids as vector
mtcars <- mtcars %>% mutate(lmresids=resids) # add to df


low <- mtcars %>% summarise(low=mean(lmresids)-sd(lmresids))#calc low limit
 # assign as variable
high <- mtcars %>% summarise(high=mean(lmresids)+sd(lmresids))
low <- dplyr::pull(low)
high <- dplyr::pull(high)                   
mtcars <- mtcars %>% mutate(sds=ifelse(lmresids>low & lmresids<high,0,1)) #create new var
kable(head(mtcars[-c(6:12)]))%>% 
  kable_styling(full_width = FALSE) %>% kable_minimal()

p2 <- hpvmpg+geom_point()+geom_smooth(method='lm',formula =formula,colour='black', size=.4,alpha=.6,se=F)+
  stat_fit_deviations(formula = formula, colour = "red")+geom_label_repel(aes(label =car_name),
    arrow = arrow(length = unit(0.02, "npc")),
    box.padding = .5,min.segment.length = 0,max.overlaps = Inf)

```
 
 
### only label extreme residuals
```{r}
formula <- y ~ poly(x, 2,raw=TRUE) # calc formula for display
dat2 <- mtcars
dat2$car_name <- ""
ix_label <- which(mtcars$sds == 1)
dat2$car_name[ix_label] <- mtcars$car_name[ix_label]
hpvmpg <- dat2 %>% ggplot(aes(x=disp,y=hp,label=car_name))

hpvmpg+geom_point(color = ifelse(dat2$car_name == "", "grey50", "red"),size = ifelse(dat2$car_name == "", 1, 3))+geom_text_repel(aes(label =car_name),
    box.padding = .55)+geom_smooth(method='lm',formula =formula,colour='black', size=.3,alpha=.6,se=F)+
  stat_fit_deviations(formula = formula, colour = "red",size=.5,alpha=.4)+
  stat_poly_eq(aes(label =  paste(stat(eq.label),
                                  stat(adj.rr.label), sep = "*\", \"*")),
               formula = formula, parse = TRUE,size=3,  geom = "label_npc")+theme_minimal()

```

## line defined by equation to scatterplot
```{r}
plot2<- mtcars %>% ggplot(aes(x=hp,y=qsec))
q1 <- plot2+geom_point()+
    stat_function(fun = function(x) 20-(.013*x)) # linear function
fun = 'y = 20 - 0.013x - 0.00003x^2'
q2 <- plot2+geom_point()+    stat_function(fun = function(x) 20-(.013*x+.00003*x^2))# poly function
 # poly function
q2 <- q2+ annotate("text", x = 175, y = 22, label = fun,size=4)
grid.arrange(q1,q2,nrow=1)
```





